---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
---

Peer Assessment 1

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", header = T, stringsAsFactors = F)
str(activity)
activity$date <- as.Date(activity$date) 
#activity$date <- as.Date(activity$date, "%Y-%m-%d") 
str(activity)
```

## What is mean total number of steps taken per day?
```{r echo=F}
#library(doBy)
#activity.doby <- summaryBy(steps~date, data=activity, FUN=c(sum))
#activity.doby
```

```{r}
library(data.table)
activity.DT = data.table(activity)
str(activity.DT)
dailyStep <- activity.DT[,list(sumSteps=sum(steps)), by=date]
```

###1. Make a histogram of the total number of steps taken each day
```{r}
hist(dailyStep$sumSteps, xlab="Total Number of Steps taken per Day", ylab="Frequency", main="Histogram")
```

###2. Calculate and report the mean and median total number of steps taken per day  

Mean total number of steps taken per day
```{r}
steps.mean <- mean(dailyStep$sumSteps, na.rm=T) #  ignore the missing values in the dataset
steps.mean
```

Median total number of steps taken per day
```{r}
steps.median <- median(dailyStep$sumSteps, na.rm=T) #  ignore the missing values in the dataset
steps.median
```

 


## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}

meaninterval <- activity.DT[,list(meanintervalSteps=mean(steps, na.rm=T)), by=interval]

plot.ts(meaninterval$interval, meaninterval$meanintervalSteps, type="l", xlab="5-minute interval", ylab="average number of steps")

```
### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
meaninterval$interval[which.max(meaninterval$meanintervalSteps)]
```


## Imputing missing values

1. the total number of rows with NAs :
```{r}
sum(is.na(activity$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

My strategy for filling is using the mean for that 5-minute interval
```{r}
newdataset <- activity
for(i in 1:nrow(newdataset)){
  if(is.na(newdataset$steps[i])){
    newdataset$steps[i] <- meaninterval[ which(meaninterval$interval==newdataset$interval[i]),]$meanintervalSteps
  }
}
```

```{r}
#original data 
head(activity)
#a new dataset that is equal to the original dataset but with the missing data filled in.
head(newdataset)
```
### histogram of the total number of steps taken each day
```{r}
library(dplyr)
newdataset.daysum <- newdataset %>% group_by(date) %>% summarize(daysum =sum(steps,na.rm = TRUE))
hist(newdataset.daysum$daysum)
```

```{r}
mean(newdataset.daysum$daysum)
median(newdataset.daysum$daysum)
```

first step of mean and median 
```{r}
steps.mean
steps.median
```

### What is the impact of imputing missing data on the estimates of the total daily number of steps?  
: imputing of missing value affect to median but not to mean. 

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=FALSE}
Sys.setlocale("LC_TIME","C")
```

```{r}
#  Use the dataset with the filled-in missing values for this part.
head(newdataset)
newdataset$weekday <- ifelse(weekdays(newdataset$date)=="Sunday"|weekdays(newdataset$date)=="Saturday", "weekend", "weekday")
newdataset$weekday <- factor(newdataset$weekday)
table(newdataset$weekday)
head(newdataset)
```

### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
newdataset.DT <- data.table(newdataset)
newmeaninterval <- newdataset.DT[,list(meanintervalSteps=mean(steps)), by=list(interval, weekday)]
newmeaninterval
```


### 2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
library(lattice)
xyplot(newmeaninterval$meanintervalSteps ~ newmeaninterval$interval | newmeaninterval$weekday, 
       layout = c(1, 2), type = "l", 
       xlab = "5-minutes interval", ylab = "the average number of steps")
```
